---
title: "R course at NHM"
subtitle: "trees"
author: "Srishti Arya"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
---

# setup
```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(stringr)
library(ape)
library(ggtree) # installed via bioconductor

knitr::opts_knit$set(root.dir = "/Users/srishti/Desktop/OneDrive - Queen Mary, University of London/SrishtiArya_AllWork/Projects/NHM_R/cuddly-funicular/")

options(scipen = 999)
```

# loading data example
the data is stored in LISTS
always use str() to examine data pre modification to understand data structure
```{r, fig.width=10, fig.height=50}
ape::read.tree(file = "input_data_trees/basic.tre")

ape::read.nexus(file = "input_data_trees/frog-tree.nex")
```

# loading test data
```{r}
mytree<-read.tree("input_data_trees/basic.tre")

mytree

str(mytree)
```

# plotting using base r
base r plot function goes through a loop
it finds that this is phylogenetic data so it uses plot.phylo from ape package
can also input function plot.phylo directly
```{r}
plot.phylo(mytree)
```

# bird data example
```{r}
bird.orders<-read.tree("input_data_trees/bird.orders.tre")

plot.phylo(bird.orders, 
           cex = 0.6,
           no.margin = TRUE)

plot.phylo(bird.orders, 
           cex = 0.6,
           type = "fan", # this is making a circle phylogeny instead of a cladogram
           no.margin = TRUE)

# the birds have qualifiers
birdfacts<-read.csv("input_data_trees/bird-facts.csv")
view(birdfacts)

bird.colours<-c("lightblue3", "salmon", "darkolivegreen3")

plot(bird.orders, 
     label.offset = 1,
     cex = 0.9,
     no.margin = TRUE)

tiplabels(pch = 22,
          bg = bird.colours[as.factor(birdfacts$coolness)],
          cex = 1.2,
          adj = 1)
```
# note on base r plotting
the code for this example only works if your tip labels are in the same order as the data in the data frame
if not, it will be random
we can reorder the data to match the phylogeny
```{r}
birdfacts <- birdfacts[match(bird.orders$tip.label,     
birdfacts$orders), ]
```

# ggtree
```{r fig.width=7, fig.height=7}
ggtree(mytree)

ggtree(bird.orders, 
       color = "blue", 
       size = 2)+
  xlim(0, 35)+
  geom_tiplab(offset = 1,
              fontface = "italic")+
  theme_tree2()+
  geom_nodepoint(colour = "coral", 
                 shape = "square", 
                 size = 5,
                 alpha = 0.8) +
  geom_tippoint(size = 2, colour = "cyan3")+ 
  geom_text(aes(label = node), 
            hjust = -0.3)

ggtree(bird.orders, layout = "fan")+
  xlim(0, 35)+
  geom_tiplab(offset = 1,
              fontface = "italic")+
  theme_tree()

```



